---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---

This is to remove the messages all over the document.
```{r setup}
knitr::opts_chunk$set(message = F, warning = F)
```


## Loading and preprocessing the data
```{r loading_reprocessing}
library(readr)
activities <- read_csv("activity.zip", col_types = cols())

library(dplyr, quietly = TRUE)
mutated <- mutate(activities, clean_steps = ifelse(is.na(steps), 0, steps))
mutated
```


## What is mean total number of steps taken per day?
```{r statistics_per_day}
grouped_dates <- group_by(mutated, date)
summarized_daily_steps <- summarise(grouped_dates, total_steps = max(clean_steps))

library(ggplot2)
qplot(summarized_daily_steps$total_steps, geom="histogram", xlab="Daily total steps") 
daily_total_mean <- mean(summarized_daily_steps$total_steps)
daily_total_median <- median(summarized_daily_steps$total_steps)
```
The mean of the total number of steps taken per day is `r daily_total_mean` and the median is `r daily_total_median`

## What is the average daily activity pattern?



```{r average_interval_steps}
grouped_intervals <- group_by(mutated, interval)
summarized_interval_average <- summarise(grouped_intervals, average_steps = mean(clean_steps))

qplot(interval, average_steps, data = summarized_interval_average, geom="path", ylab="Average steps") 

max_average_steps <- max(summarized_interval_average$average_steps)
filtered_max_steps <- filter(summarized_interval_average, average_steps == max_average_steps)
max_average_interval <- unique(filtered_max_steps$interval)
```
The 5-minute interval, on average across all the days in the data set, contains the maximum number of steps is `r max_average_interval`

## Imputing missing values



```{r missing_total_number }
missing_values <- filter(mutated, is.na(steps)) 
total_missing_values <- count(missing_values)
```
The total number of missing values in the data set is `r total_missing_values`
```{r imputing-missing_values}
imputed <- left_join(missing_values, summarized_interval_average, by = "interval")
imputed_join_columns <- select(imputed, date, interval, average_steps)
mutated_imputed <- 
  left_join(mutated, imputed_join_columns, by = c("date", "interval"))
mutated_imputed_unified <- mutate(mutated_imputed, imputed_steps = ifelse(is.na(steps), average_steps, steps))

grouped_imputed_dates <- group_by(mutated_imputed_unified, date)
summarized_imputed_steps <- summarise(grouped_imputed_dates, total_steps = max(imputed_steps))

library(ggplot2)
qplot(summarized_imputed_steps$total_steps, geom="histogram", xlab="Daily imputed total steps") 
imputed_total_mean <- mean(summarized_imputed_steps$total_steps)
imputed_total_median <- median(summarized_imputed_steps$total_steps)
```
The mean of the total number of steps taken per day after imputation is `r imputed_total_mean` and before imputation was `r daily_total_mean` and the median is `r imputed_total_median` and before imputation was `r daily_total_median`, which indicates almost no change in the reported measures, where mean is slightly increased but median didn't change at all, so, one can say that imputing the missing values didn't impact the original estimates of the total daily number of steps.

## Are there differences in activity patterns between weekdays and weekends?

```{r activity_patterns}
library(chron)
day_type_unified <- mutate(mutated_imputed_unified, day_type = factor( ifelse(is.weekend(date), "weekend", "weekday")))

unified_grouped_intervals <- group_by(day_type_unified, day_type, interval)
summarized_unified_average <- summarise(unified_grouped_intervals, average_steps = mean(imputed_steps))

qplot(interval, average_steps, data = summarized_unified_average, geom="path", ylab="Average steps", facets = "day_type")
```